---
title: "NBA_KNN_Predictions"
author: "Ryan Salsbury"
date: "10/16/2018"
output: html_document
---
Predict Offensive and Defensive Win Shares for each NBA Player for the 2018 Season using a K-Nearest Neighbor Model. All Data preparation was done in excel. The two input files include cumulative stats up until each season of a player's career and the following season's OWS & DWS (objective variables). The goal was to find the most similar players historically to predict a player's OWS and DWS for the upcoming 2018-19 season. 

Load Packages and read data
```{r}
library(caTools)
library(caret)
library(sqldf)

OWS <- read.csv("KNN_OWS.csv")
DWS <- read.csv("KNN_DWS.csv")
```

Create OWS Model
```{r}
#split into testing and training sets with 95% data in training
set.seed(123)
split = sample.split(OWS$OWS_Next, SplitRatio = 0.95)
training_set = subset(OWS, split == TRUE)
test_set = subset(OWS, split == FALSE)

#remove player names
training = training_set[,-1]
test = test_set[,-1]

#remove objective variable
x.train = training[,-17]
y.train = training[,17]
x.test = test[,-17]
y.test = test[,17]

#setup cross validation with 10 folds
trControl <- trainControl(method  = "cv",
                          number  = 10)

#train model to find most optimal K nearest neighbors(1-50) using RMSE as evaluator
OWS <- train(OWS_Next ~ .,
             method     = "knn",
             tuneGrid   = expand.grid(k = 1:50),
             trControl  = trControl,
             metric     = "RMSE",
             data       = training)

#viewing the OWS variable shows that the optimal number of k is 18. 
print(OWS)

#validate on test data
OWS_Predict <- predict(OWS,newdata = x.test)
sqrt(mean((OWS_Predict - y.test)^2))

```

Create DWS Model
```{r}
#split into testing and training sets with 95% data in training
set.seed(123)
split = sample.split(DWS$DWS_Next, SplitRatio = 0.95)
training_set = subset(DWS, split == TRUE)
test_set = subset(DWS, split == FALSE)

#remove Player Names
training = training_set[,-1]
test = test_set[,-1]

#remove objective variable
x.train = training[,-13]
y.train = training[,13]
x.test = test[,-13]
y.test = test[,13]

#setup cross validation with 10 folds
trControl <- trainControl(method  = "cv",
                          number  = 10)

#train model to find most optimal K nearest neighbors(1-50) using RMSE as evaluator
DWS <- train(DWS_Next ~ .,
             method     = "knn",
             tuneGrid   = expand.grid(k = 1:50),
             trControl  = trControl,
             metric     = "RMSE",
             data       = training)

#viewing the DWS variable shows that the optimal number of k is 20. 
print(DWS)

#validate on test data
DWS_Predict <- predict(DWS,newdata = x.test)
sqrt(mean((DWS_Predict - y.test)^2))

```

Make Predictions for this year
```{r}

#load new data for players this season
Predict_OWS <- read.csv("Predict_OWS.csv")
Predict_DWS <- read.csv("Predict_DWS.csv")

#remove player names
Predict_OWS_data = Predict_OWS[,-1]
Predict_DWS_data = Predict_DWS[,-1]

#run model on new data
OWS_Predict <- predict(OWS,newdata = Predict_OWS_data)
DWS_Predict <- predict(DWS,newdata = Predict_DWS_data)

#convert predictions to data frame
OWS_Predict <- as.data.frame(OWS_Predict)
DWS_Predict <- as.data.frame(DWS_Predict)

#join results to player names
Predictions <- sqldf("select po.Player, o.OWS_Predict, d.DWS_Predict from Predict_OWS po JOIN OWS_Predict o ON po.rowid = o.rowid JOIN DWS_Predict d ON po.rowid = d.rowid")


#view results
print(Predictions)

#export results to csv
write.csv(x = Predictions, file = "NBA_WS_Projections.csv")
```